The Effects of the NBA COVID Bubble on the NBA Playoffs

### 2020-11-13 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

Tests for differences in proportions were conducted for comparing winning percentage
and the three different shooting percentages in 2020 to those of
2017-2019. The null hypothesis is 
\[
H_0: p_1 = p_2,
\]
where $p_1$ and $p_2$ are the shooting percentage in 2020 and
2017-2019, respectively. Or $p_1$ and $p_2$ are the home winning percentage for 2020 and
2017-2019, respectively. This proportion comes from summing the number of home teams wins
and dividing by the total number of games played.
$\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
proportion for 2017-2019. The test statistic is
where $\hat p$ is the pooled shooting percentage of 2017-2020. Or where $\hat p$ is the 
pooled home team winning percentage of 2017-2020.
The null distribution of the test statistic is also standard normal.
The test will be conducted seven times, once comparing home winning percentage for 
2020 to that of 2017-2019 and once for each of two-point shooting, three-point shooting and
free throws for both home and away teams in 2020 vs. 2017-2019. These hypotheses will also be
tested using both a z-test and a Wilcoxon test.

# Results {#sec:results}
```{r, echo=FALSE}
#Analysis
one <- prop.test(x=c(sum(game_log_2020_home$W.L), sum(game_log_2017_2019_home$W.L)),
                 n=c(nrow(game_log_2020_home), nrow(game_log_2017_2019_home)))
onew <- wilcox.test(x = game_log_2020_home$W.L, y = game_log_2017_2019_home$W.L)
rowone <- c(as.numeric(one$estimate[1]), as.numeric(one$estimate[2]), one$p.value, onew$p.value)

@@

titles <- c(2020, "2017-19", "Z-test P-value", "Wilcoxon P-value")
rows <- c("Home Win ", "Home Scoring", "Away Scoring", "Home 2P", "Home 3P",
          "Home FT", "Away 2P", "Away 3P", "Away FT")
results <- data.frame(rowone, rowtwo, rowthree, rowfour, rowfive, rowsix,
                    rowseven, roweight, rownine, stringsAsFactors = FALSE)
resultsfinal <- data.frame(t(results))
resultsfinal1 <- round(resultsfinal[1:2], 3)
resultsfinal2 <- round(resultsfinal[3:4], 4)
resultsfinal <-cbind(resultsfinal1, resultsfinal2)
colnames(resultsfinal) <- titles
row.names(resultsfinal) <- rows
```

@@

First, we see a statistically significant change
in home win percentage in 2020 from 2017-19, with p-value of 0.0365 for the
z-test and 0.0369 for the Wilcoxon test. The 95% confidence interval of (-0.255, -0.008) 
confirms our belief that home-court advantage was lost in the 2020 NBA playoffs. However, after accounting for
multiple tests using the Bonferroni correction, the p-values for both tests are no longer significant.
So, we may only cautiously say there is evidence that home-court advantage was
not a factor in 2020.

>>>>

Tests for differences in proportions were conducted for comparing winning percentage
and the three different shooting percentages in 2020 to those of
2017-2019. The null hypothesis is 
\[
H_0: p_1 = p_2,
\]
where $p_1$ and $p_2$ are the shooting percentage in 2020 and
2017-2019, respectively. Or $p_1$ and $p_2$ are the home winning percentage for 2020 and
2017-2019, respectively. This proportion comes from summing the number of home teams wins
and dividing by the total number of games played.
$\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
proportion for 2017-2019. The test statistic is
where $\hat p$ is the pooled shooting percentage of 2017-2020. Or where $\hat p$ is the 
pooled home team winning percentage of 2017-2020.
The null distribution of the test statistic is also standard normal.
The test will be conducted seven times, once comparing home winning percentage for 
2020 to that of 2017-2019 and once for each of two-point shooting, three-point shooting and
free throws for both home and away teams in 2020 vs. 2017-2019. These hypotheses will also be
tested using both a z-test and a nonparametric test. The Wilcoxon test will be used for shooting percentages and
Fisher's test will be used to test the binary variable of home team winning.

# Results {#sec:results}
```{r, echo=FALSE}
#Analysis
one <- prop.test(x=c(sum(game_log_2020_home$W.L), sum(game_log_2017_2019_home$W.L)),
                 n=c(nrow(game_log_2020_home), nrow(game_log_2017_2019_home)))
X2020 <- table(game_log_2020_home$W.L)
X17_19 <- table(game_log_2017_2019_home$W.L)
homewin<- rbind(X2020, X17_19)
onew <- wilcox.test(x = game_log_2020_home$W.L, y = game_log_2017_2019_home$W.L)
rowone <- c(as.numeric(one$estimate[1]), as.numeric(one$estimate[2]), one$p.value, onew$p.value)

@@

titles <- c(2020, "2017-19", "Z-test P-value", "Nonparametric P-value")
rows <- c("Home Win ", "Home Scoring", "Away Scoring", "Home 2P", "Home 3P",
          "Home FT", "Away 2P", "Away 3P", "Away FT")
results <- data.frame(rowone, rowtwo, rowthree, rowfour, rowfive, rowsix,
                    rowseven, roweight, rownine, stringsAsFactors = FALSE)
resultsfinal <- data.frame(t(results))
resultsfinal1 <- round(resultsfinal[1:2], 3)
resultsfinal2 <- round(resultsfinal[3:4], 4)
resultsfinal <-cbind(resultsfinal1, resultsfinal2)
colnames(resultsfinal) <- titles
row.names(resultsfinal) <- rows
```

@@

First, we see a statistically significant change
in home win percentage in 2020 from 2017-19, with p-value of 0.0365 for the
z-test and 0.0400 for Fisher's test. The 95% confidence interval of (-0.255, -0.008) 
confirms our belief that home-court advantage was lost in the 2020 NBA playoffs. However, after accounting for
multiple tests using the Bonferroni correction, the p-values for both tests are no longer significant.
So, we may only cautiously say there is evidence that home-court advantage was
not a factor in 2020.
